Collaborative Infrastructure for Test-Driven\\Scientific Model Validation

### 2014-03-14 sciunit-icse14/sciunit-icse14.tex
ICSE14: Revised code, sec1+2, entering sec3+4

We argue that a similar test-driven methodology would be valuable to scientific communities as they seek to  validate increasingly complex models against growing repositories of empirical data. Scientific communities differ from software communities in several key ways, however. In this paper, we introduce \emph{SciUnit}, a framework for test-driven scientific model validation and outline how SciUnit, supported by new and existing collaborative infrastructure, could be  integrated into the modern scientific workflow.
%in the context of the applications and architectures that the language may need to target.
%
%Programming languages targeted toward end-users in high-performance computing must consider a number of issues in addition to performance. Typically, researchers have focused verifiability, portability and ease-of-use, and the right balance often depends on the particular uses that they are put toward.
%
%Ace %, evaluates it against several design and adoption criteria, and describes several use cases as well as a realistic case study involving parallelizing neurobiological circuit simulations on clusters of GPUs. 
%is an extensible statically-typed programming language embedded within the ubiquitous scripting language Python. Python serves as a {metalanguage}, enabling programmatic control over several aspects of language specification and compilation. 
%We demonstrate the power of this design for HPC by defining all of the primitives of the OpenCL kernel language as a library of user-defined types in Ace, then extending this core with several higher-level first-class abstractions. The target of translation, called the \emph{backend}, is user-defined as well, allowing us to implement the semantics of these primitives by direct translation to OpenCL, CUDA or C99.
%
%Although Ace is built around a static type discipline, it supports a novel form of generic programming based on type propagation that, together with a form of extensible local type inference, eliminates most of the syntactic overhead usually associated with statically-typed languages. A companion host API allows type information from \verb|numpy| arrays to directly propagate into Ace kernels, supporting seamless adoption of Ace into current scientific workflows.
%
\end{abstract}
\category{D.2.5}{Software Engineering}{Testing and Debugging}%[complexity measures, performance measures]
\terms{Documentation, measurement, standardization}
\keywords{unit testing, model validation, cyberinfrastructure}

@@

\section{Introduction}
\begin{figure*}[t]
\vspace{-30px}
\centering
\includegraphics[scale=0.45]{diagram1.pdf}
\vspace{-40px}
\caption{\small Tests are derived from data and models are derived from scientific theories. The score table summarizes the performance of a collection of models against a suite of tests. A table is generated and visualized in an IPython notebook stored inside a \emph{suite repository} (e.g. \texttt{saturnsuite}) hosted on social coding infrastructure (here, Github). SciDash is a portal that discovers and organizes these suite repositories and provides read-only views of the notebooks they contain. Common testing code, capabilities and bridges to external model and data repositories are also collaboratively developed in \emph{common repositories} (e.g. \texttt{cosmounit}).}  
\label{fig:sciunit_overview}
\vspace{-10pt}
\end{figure*}

@@

Quantitative models are today validated by \emph{peer review}. For a model to be accepted by a scientific community, its advocates must submit a paper describing how it works and providing evidence that it predicts a quantity of interest more accurately than previous models, or in some cases, that it makes a desirable tradeoff  between accuracy and complexity \cite{box1987empirical}. Other members of the relevant community are then tasked with ensuring that validity was measured properly and that relevant data and competing models were adequately considered, drawing on knowledge of statistical methods and on the prior literature. Publishing is a primary motivator for most scientists \cite{howison2011scientific}.

Quantitative scientific modeling and software development share much in common. Indeed, quantitative models are increasingly being implemented in software and in some cases, the software \emph{is} the model (e.g. complex simulations). The peer review process for papers is similar in many ways to the \emph{code review} process used in many development teams, where team members look for mistakes, enforce style and architectural guidelines and check that the code is \emph{valid} (i.e. that it achieves its intended goal) before permitting it to be committed to the primary source code repository.

Code review can be quite effective \cite{codereview}, but this requires that developers  expend considerable effort \cite{kemerer2009impact}. 
%Code review is also most effective for resolving  issues related to software architecture. 
Most large development teams thus supplement code reviews with more {automated} approaches to verification and validation, the most widely-used  of which is \emph{unit testing} \cite{beck2003}. In brief, unit tests are functions that check that the behavior of a single component satisfies a single functional  criterion. 
%A suite of such tests can be considered a partial specification of the program or component under development.  
A suite of such tests complements code review by making it easier to answer questions like these more easily:
\begin{enumerate}
\item What functionality should each component have?
\item What functionality has been adequately implemented? What remains to be done?
%\item How much progress does an individual code contribution make? 
\item Does a candidate code contribution cause \emph{regressions} in other parts of a program?
\end{enumerate}
Scientists ask analagous questions:
\begin{enumerate}
\item What are the contemporary community standards for measuring goodness-of-fit?
\item Which observations are already explained by existing models? What are the best models? What are the open modeling problems of interest?
%\item How much better is a new model than the state of the art? What are the open problems in my field?
\item How do newly-made experimental observations impact the validity of previously-published models? Can new models explain previously published data?
\end{enumerate}
But while software engineers can rely on a program's test suite, scientists today must extract this information from a body of scientific publications. This is increasingly difficult. Each publication focuses on just one model and is frozen in time, so it does not consider the latest experimental data or statistical methods. Discovering, precisely characterizing and comparing models to discover the state of the art and find open modeling problems can require an encyclopedic knowledge of the literature, as can finding all data relevant to a model. Senior scientists often attempt to fill this need by publishing review papers, but in many areas, the number of publications generated every year can be overwhelming \cite{jinha_article_2010}, and comprehensive reviews of a particular area are published relatively infrequently. Statisticians often complain that scientists are not following best practices and that community standards evolve too slowly because a canonical paper or popular review used outdated methods. Furthermore, if the literature simply doesn't address an important question of validity, because  new data has been gathered since the model paper was published for example, a researcher might need to reimplement an existing model in order to evaluate it.

One might compare this to a large software development team answering the questions listed above based only on carefully reviewed component documentation together with annual high-level project summaries.
Although certainly a caricature, this motivates our suggestion that the scientific process could be improved by the adoption of test-driven methodologies alongside traditional peer review. However, the scientific community presents several unique challenges that must be addressed by any such methodology:
\begin{enumerate}
\item Unit tests are typically pass/fail, while goodness-of-fit between a model and data is typically measured by a continuous metric (e.g. a $p$-value). 
\item Unit tests often test a \emph{particular} component, whereas a \emph{validation test} must be able to handle and compare many models capable of predicting the same quantity. %They are sometimes reluctant to share certain details of their implementations \cite{badscientists}.
%\item Experimental data can itself be erroneous, noisy or inconsistent with other available data.
\item Different modelers use different programming languages.
\item Professional software developers are typically trained in testing practices and tools, while scientists  rarely have training or experience with testing practices \cite{oai:open.ac.uk.OAI2:17673}. Thus, the framework must be as simple as possible.
\item Different communities, groups and individuals prefer different goodness-of-fit metrics and care about different sets of observable quantities. In contrast, there is typically more pressure to agree upon requirements and priorities in a software development project.
\end{enumerate}

To address challenges 1-4, we will introduce a lightweight scientific validation testing framework, \textit{SciUnit}, in Sec. 2. Challenge 5 has to do with coordination between scientists. To address this, we introduce a community workflow based on widely-adopted social coding tools (here Github) along with a lightweight community portal called \textit{SciDash}. The overall goal of this work is to help scientists generate and examine tables like the one central to Figure 1, where the relative validity of a set of models having a common scope can be determined by examining scores produced by a suite of validation tests constructed from experimental data.

%Anecdotally, one of the greatest fears of new scientists is that a problem that they are working on has already been solved and published.

@@

% Section 2.
\section{Validation Testing with {SciUnit}}
\begin{figure}[t]
\small
\begin{python}
class PositionTest(sciunit.Test):
    """Tests a planetary position model based on positions observed on day n given the positions in the n-1 previous days.
    Metric: Standard p-value.
    Parameters: 
      obs_histories : list[list[position]]
      obs_positions: list[position]"""
	def __init__(self, obs_histories, obs_positions):
		self.obs_histories = obs_histories
		self.obs_positions = obs_positions

	required_capabilities = [PredictsPlanetaryPosition]

	def _judge(self, model):
		predictions = []
		for obs_history in self.obs_histories:
			predictions.append(model.predict_next_pos(obs_history))
		p = pooled_p_val(predictions, self.obs_positions)
		return sciunit.PValue(p, related_data={
			'obs_histories': obs_histories,
			'obs_positions': obs_positions,
			'predictions': predictions
		})
\end{python}
\vspace{-5px}
\caption{An example test class in \texttt{cosmounit}.}
\label{fig:rate_test}
\vspace{-15px}
\end{figure}

@@

Before explaining the details of this example, we point out that SciUnit is implemented in Python. Python is one of the most widely used languages in science today \cite{sanner1999python}. It supports calling into  many of the other popular languages, including R, MATLAB, C and Java, more cleanly than these languages support calling into Python (challenge 3). It is widely-recognized as being easy-to-read and its object system can be used to define abstract interfaces, which we leverage in support of challenge 2.

A \textit{SciUnit} validation test is an {instance} of a Python class implementing the \verbx{sciunit.Test} abstract interface (line 1). 
Here, the class \verbx{PositionTest} takes two \emph{parameters} in its constructor (constructors are named \verbx{\_\_init\_\_} in Python, lines 7-9). 
The meaning of each parameter along with a description of the goodness-of-fit metric used by the test is documented on lines 2-6. 
%For convenience, we also make use of functions provided by the popular NumPy\cite{numpy_url} and SciPy\cite{scipy_url} libraries, although these are not required by \textit{SciUnit}.  
To create a \emph{particular} position test, we instantiate this class with particular planetary observations. 
For example, the subset of cosmologists interested specifically in Saturn might instantiate a test by randomly chunking observations made about Saturn as follows:
\begin{python}
  h, p = randomly_chunk(saturn_obs_positions)
  saturn_position_test = PositionTest(h, p)
\end{python}

The class \verb|PositionTest| defines logic that is not specific to any particular planet, so it is contained in a package shared by all cosmologists called \verbx{cosmounit}, while the particular test above would be used in a test suite focused specifically on Saturn called \verbx{saturnsuite}. Both the common logic and specific suites are collaboratively developed by these overlapping research communities in source code repositories on Github (or another similar service).

Classes that implement the \verbx{sciunit.Test} interface must contain a \verbx{_judge} method that receives a candidate \emph{model} as input and produces a \textit{score} as output. 
To specify the interface between the test and the model, the test author provides a list of \emph{capabilities} in the \verbx{required_capabilities} attribute, seen on line 11 of Fig. \ref{fig:rate_test}. 
Capabilities are simply collections of methods that a test will need to invoke in order to receive relevant data, and are analogous to \emph{interfaces} in e.g. Java. 
In Python, capabilities must be written as classes with unimplemented members. 
The capability required by the test in Figure \ref{fig:rate_test} is shown in Figure \ref{fig:capability}. 
In \textit{SciUnit}, classes defining capabilities are tagged as such by inheriting from \verbx{sciunit.Capability}. The test in Figure \ref{fig:rate_test} repeatedly uses this capability on line 16 to produce a position prediction for each observation.  We assume that positions are represented in a standardized manner specified within \verbx{cosmounit}. A model is simply an object that implements capabilities (via Python's simple inheritance mechanism).

The remainder of the \verbx{_judge} method compares the model predictions to observed data to produce a pooled $p$ value. The method returns an instance of \verbx{sciunit.PValue}, a subclass of \texttt{sciunit.Score} that has been included with \textit{SciUnit} due to its generality. In addition to the $p$-value itself, the returned score object also contains metadata, via the \verbx{related_data} parameter, for  scientists who may wish to examine the result in more detail later. This illustrates some key differences between unit testing, which would simply produce a boolean result, and our conception of scientific validation testing (challenge 1). A score must induce an ordering, so that the table shown in Figure 1 can be sorted along its columns, and it can optionally specify a normalization scheme so the cells can be color-coded (not shown). A test can be manually executed using the \verbx{judge} method:
\begin{python}
score = saturn_position_test.judge(kepler_sat)
\end{python}

This method proceeds by first checking that the provided model implements all required capabilities before calling the test's \verbx{_judge} method to produce a score. A reference to the test and model are added to the score for convenience (accessible via the \verbx{test} and \verbx{model} attributes, respectively).

Alongside this position test, we could also provide a number of other test classes in \verbx{cosmounit} and instantiate them with data about aspects of Saturn's motion (e.g. it's velocity)  to produce a comprehensive suite in \verbx{saturnsuite}. 
\begin{python}
saturn_motion = sciunit.TestSuite([saturn_position_test, saturn_velocity_test, ...])
\end{python}
Like a single test, a test suite is capable of judging one or more models. The result is a score matrix much like the one diagramed in Fig. \ref{fig:sciunit_overview}.
\begin{python}
sm_matrix = saturn_motion.judge([copernicus_sat, ptolemy_sat, newton_sat, kepler_sat])
\end{python}
A test suite requires the union of the capabilities required by the tests it contains. A model that performs well across tests in such a suite could reasonably claim (e.g. to reviewers) that it is a coherent, valid model of Saturn's motion. As new data is collected, new tests can be added to the suite. However, because the interface between the test and the model remains the same (only the data parameterizing the tests changes), the table can be updated automatically. Similarly, when a new model is developed, it can immediately be evaluated against all known data that has been encoded as a test by simply exposing its predictions via the capabilities the test requires. %They then defined a combined metric favoring models that broadly succeeded at meeting these criteria, to produce an overall ranking. 
%Such combined criteria are simply validation tests that invoke other tests to produce a result.

@@

\begin{figure}
\begin{python}
class PredictsPlanetaryPosition(sciunit.Capability):
  def predict_next_pos(self, history): 
    """Takes a list of previous positions and produces the next position."""
    raise NotImplementedError("Model does not implement capability.")
\end{python}
\caption{An example capability specifying a single required method (used by the test in Figure \ref{fig:rate_test}).}
\label{fig:capability}
\vspace{-10px}
\end{figure}

@@

%\subsection{Models}
%Capabilities are \emph{implemented} by models. In \textit{SciUnit}, models are instances of Python classes that inherit from \verbx{sciunit.Model}. Like tests, the class itself represents a family of models, parameterized by the arguments of the constructor. A particular model is an instance of such a class.
%
%Figure \ref{fig:simple_model} shows how to write a simple {family} of models, \verbx{LinearModel}, that implement the capability in Fig. \ref{fig:capability} as well as another capability shown in Fig. \ref{fig:training}, which we will discuss below. 
%Models in this family generate a spike count by applying a linear transformation to the mean of the provided input current. The family is parameterized by the scale factor and the offset of the transformation, both scalars. 
%To create a \emph{particular} linear model, a modeler can provide particular parameter values, just as with test families:
%\begin{python}
%CA1_linear_model_heuristic = LinearModel(3.0, 1.0)
%\end{python}
%Here, the parameters to the model were picked by the modeler heuristically, or based on externally-available knowledge. 
%An alternative test design would add a training phase where these parameters were fit to data using the capability shown in Fig. \ref{fig:training}. 
%This test could thus only be used for those models for which parameters can be adjusted without human involvement. 
%Whether to build a training phase into the test protocol is a choice left to each test development community. 
%
%Fig. \ref{fig:rate_test} does not include a training phase. 
%If training data is externally available, models that nevertheless do implement a training capability (like \verb|LinearModel|) can simply be trained explicitly by calling the capability method just like any other Python method:
%\begin{python}
%CA1_linear_model_fit = LinearModel()
%CA1_linear_model_fit.train_with_currents(CA1_training_in, CA1_training_out)
%\end{python}
%\begin{figure}
%\begin{python}
%class LinearModel(sciunit.Model, SpikeCountFromCurrent, 
%    TrainSpikeCountFromCurrent):
%  def __init__(self, scale=None, offset=None): 
%    self.scale, self.offset = scale, offset
%    
%  def spike_count_from_current(self, input):
%    return int(self.scale*numpy.mean(input) + self.offset)
%
%  def train_with_currents(self, currents, counts):
%    means = [numpy.mean(c) for c in currents]
%    [self.offset, self.scale] = numpy.polyfit(means, counts, deg=1)    
%\end{python}
%\caption{A model that returns a spike count by applying a linear transformation to the mean input current. The parameters can be provided manually or learned from data provided by a test or user (see text).}
%\label{fig:simple_model}
%\vspace{-15px}
%\end{figure}
%
\section{Collaborative Workflow}
The design just described has been purposefully left as simple as possible, in pursuit of challenge 4. Despite its simple design, it captures the essential elements of the scientific model validation process and satisfies the four challenges we laid out. SciUnit can be used by individual scientists to organize their workflows, but because science, and in particular, model validation is a collaborative process, we have also described the intended use of SciUnit within a collaborative workflow, mediated by a social coding tool like Github.

@@

\section{Discussion}
Academic peer review and code review are similar, but software developers typically supplement human review with test-driven methodologies. Such methodologies may benefit scientific communities as well, but several unique  constraints make reusing existing testing frameworks and processes difficult. We describe a core testing framework that addresses these issues  by using an existing, widely-adopted language and designing a flexible, simple framework that captures the domain-specific structure of scientific model validation. We then describe, by example, how the various components can be organized into software repositories and collaboratively maintained using social coding tools to support existing scientific practices. We end by outlining a collaboratively filtered portal that sits atop these tools and facilitates exploratory analyses and repository discovery.

>>>>

We argue that a similar test-driven methodology would be valuable to scientific communities as they seek to  validate increasingly complex models against growing repositories of empirical data. Scientific communities differ from software communities in several key ways, however. In this paper, we introduce \emph{SciUnit}, a framework for test-driven scientific model validation and outline how SciUnit, supported by new and existing collaborative infrastructure, could be  integrated into the modern scientific workflow.
\end{abstract}
\category{D.2.5}{Software Engineering}{Testing and Debugging}%[complexity measures, performance measures]
\terms{Documentation; measurement; standardization}
\keywords{unit testing; model validation; cyberinfrastructure}

@@

\section{Introduction}
\begin{figure*}[t]
\vspace{-30px}
\centering
\includegraphics[scale=0.45]{diagram1.pdf}
\vspace{-50px}
\caption{\small Tests are derived from data and models are derived from scientific theories, either directly or through existing data and model repositories. The score table summarizes the performance of a collection of models against a suite of tests. Tests and models are initialized, executed and visualized using \texttt{sciunit} in an IPython notebook stored inside a \emph{suite repository} (e.g. \texttt{saturnsuite}) hosted on a social coding website (e.g. Github). SciDash  discovers and organizes  suite repositories and provides interactive views of the notebooks they contain. Common testing code, capabilities and bridges to external model and data repositories are also collaboratively developed (e.g. \texttt{cosmounit}).}  
\label{fig:sciunit_overview}
\vspace{-10pt}
\end{figure*}

@@

Quantitative models are today validated by \emph{peer review}. For a model to be accepted by a scientific community, its advocates must submit a paper that describes how it works and provides evidence that it makes more accurate predictions than previous models (or that it makes a desirable tradeoff  between accuracy and complexity) \cite{box1987empirical}. Other members of the relevant community are then tasked with ensuring that validity was measured properly and that all relevant data and competing models were adequately considered, drawing on knowledge of statistical methods and the prior literature. Publishing is a primary motivator for most scientists \cite{howison2011scientific}.

Quantitative scientific modeling shares much in common with software development. Indeed, quantitative models are increasingly being implemented in software and in some cases, the software \emph{is} the model (e.g. complex simulations). The peer review process for papers is similar in many ways to the \emph{code review} process used in many development teams, where team members look for mistakes, enforce style and architectural guidelines and check that the code is \emph{valid} (i.e. that it achieves its intended goal) before permitting it to be committed to the primary source code repository.

Code review can be quite effective \cite{codereview}, but this requires that developers  expend considerable effort \cite{kemerer2009impact}. 
%Code review is also most effective for resolving  issues related to software architecture. 
Most large development teams thus supplement code reviews with more {automated} approaches to verification and validation, the most widely-used  of which is \emph{unit testing} \cite{beck2003}. In brief, unit tests are functions that check that the behavior of a single component satisfies a single functional  criterion. 
%A suite of such tests can be considered a partial specification of the program or component under development.  
A suite of such tests complements code review by making it easier to answer questions like these:
\begin{enumerate}
\item What functionality should each component have?
\item What functionality has been adequately implemented? What remains to be done?
%\item How much progress does an individual code contribution make? 
\item Does a candidate code contribution cause \emph{regressions} in other parts of a program?
\end{enumerate}
Scientists ask analagous questions:
\begin{enumerate}
\item What are the contemporary community standards for measuring goodness-of-fit?
\item Which observations are already explained by existing models? What are the best models for a particular quantity? What are the open modeling problems of interest? What data has yet to be explained?
%\item How much better is a new model than the state of the art? What are the open problems in my field?
\item How do newly-made experimental observations impact the validity of previously published models? Can new models explain previously published data?
\end{enumerate}
But while software engineers can rely on a program's test suite, scientists today must extract this information from a body of scientific papers. This is increasingly difficult. Each paper is narrowly focused, often considering just one model or dataset, and is frozen in time, so it does not consider the latest experimental data or statistical methods. Discovering, precisely characterizing and comparing models to discover the state of the art and find open modeling problems can require an encyclopedic knowledge of the literature, as can finding all data relevant to a model. Senior scientists often attempt to fill this need by publishing review papers, but in many areas, the number of publications generated every year can be overwhelming \cite{jinha_article_2010}, and comprehensive reviews of a particular area are published infrequently. Statisticians often complain that scientists are not following best practices and that community standards evolve too slowly because a canonical paper or popular review used outdated methods. Furthermore, if the literature simply doesn't address an important question of validity, because  new data has been gathered since the model paper was published for example, a researcher might need to reimplement an existing model in order to evaluate it.

One might compare this to a large software development team answering the questions listed above based largely on carefully reviewed, but rarely updated,  API documentation and annual high-level project summaries.
While certainly a caricature, this motivates our suggestion that the scientific process could be improved by the adoption of test-driven methodologies alongside traditional peer review. However, the scientific community presents several unique challenges that must be addressed by any such methodology:
\begin{enumerate}
\item Unit tests are typically pass/fail, while goodness-of-fit between a model and data is typically measured by a continuous metric. 
\item Unit tests often test a \emph{particular} component, whereas a \emph{validation testing} system must be able to handle and compare many models with the same scope, developed by teams who only loosely coordinate their efforts. %They are sometimes reluctant to share certain details of their implementations \cite{badscientists}.
%\item Experimental data can itself be erroneous, noisy or inconsistent with other available data.
\item Scientists use a variety of programming languages.
\item Professional software developers are typically trained in testing practices and tools, while scientists  rarely have training or experience with testing practices \cite{oai:open.ac.uk.OAI2:17673}. %Thus, the framework must be as simple as possible.
\item Different communities, groups and individuals prefer different goodness-of-fit metrics and focus on different sets of observable quantities. In contrast, there is typically more pressure to agree upon requirements and priorities in a software development project.
\end{enumerate}

To begin to address these challenges, we will introduce a lightweight and flexible validation testing and framework called \textit{SciUnit} in Sec. 2. Many of these challenges have to do with coordination between scientists. To begin to address this, we then describe a community workflow based on widely-adopted social coding tools (here, Github) and a lightweight community portal called \textit{SciDash}, in Sec. \ref{scidash}. The overall goal of this work is to help scientists generate and examine tables like the one central to Figure 1, where the relative validity of a set of models having a common scope can be determined by examining scores produced by a suite of validation tests constructed from experimental data. 
We discuss adoption strategies and directions for future research into scientific model validation practices in Sec. \ref{discussion}.
%Anecdotally, one of the greatest fears of new scientists is that a problem that they are working on has already been solved and published.

@@

% Section 2.
\section{Validation Testing with {SciUnit}}

@@

Before explaining the details of this example, we point out that SciUnit is implemented in Python. Python is one of the most widely used languages in science today \cite{sanner1999python}, and is increasingly regarded as the \textit{de facto} language of open source scientific computing. It is easy to read and has a simple object system that we use to specify the interface between tests and models (challenge 2). It supports calling into  many other popular languages, including R, MATLAB, C and Java, often more cleanly than they support calling into Python (challenge 3). The IPython notebook is a powerful web-based ``read-eval-visualize loop'' that we use to support interactive table visualization, and it also permits using other languages on a per-cell basis \cite{ipynb}. Together, this makes writing wrappers around tests and models written in other languages relatively simple. We anticipate leveraging this infrastructure to make generating such wrappers nearly automatic as future work.
\begin{figure}[t]
\small
%class PositionTest(sciunit.Test):
%    """Tests a planetary position model based on positions observed on night n given the positions in the n-1 previous nights.
%    Metric: Standard p-value.
%    Parameters: 
%      obs_histories : list[list[position]]
%      obs_positions: list[position]"""
%	def __init__(self, obs_histories, obs_positions):
%		self.obs_histories = obs_histories
%		self.obs_positions = obs_positions
%
%	required_capabilities = [PredictsPlanetaryPosition]
%
%	def _judge(self, model):
%		predictions = []
%		for obs_history in self.obs_histories:
%			predictions.append(model.predict_next_pos(obs_history))
%		p = pooled_p_val(predictions, self.obs_positions)
%		return sciunit.PValue(p, related_data={
%			'obs_histories': obs_histories,
%			'obs_positions': obs_positions,
%			'predictions': predictions
%		})
\begin{python}
class PositionTest(sciunit.Test):
  """Tests a planetary position model based on positions observed on night n given the positions in the n-1 previous nights.
  Observation: {
      'histories':list[list[Position]], 
      'positions':list[Position]]
  }"""
  required_capabilities = [PredictsPlanetaryPosition]
  def generate_prediction(self, model):
    return [model.predict_next_pos(obs_history)
      for obs_history 
      in self.observation['histories']]

  def compute_score(self, observation, prediction):
    obs_positions = observation['positions']
    return pooled_p_val([abs_distance(obs, pred)
      for (obs, pred) 
      in zip(obs_positions, prediction)])
\end{python}
\vspace{-12px}
\caption{\small An example test class in \texttt{cosmounit}.}
\label{fig:rate_test}
\vspace{-15px}
\end{figure}

A \textit{SciUnit} validation test is an {instance} of a Python class implementing the \verbx{sciunit.Test} interface (line 1). 
A test takes one required parameter in its constructor (inherited from the base class) containing the observations against which the test will validate models.
For the class \verbx{PositionTest}, this observation parameter is a dictionary containing two types of observations (documented per Python conventions on lines 2-6). 
%For convenience, we also make use of functions provided by the popular NumPy\cite{numpy_url} and SciPy\cite{scipy_url} libraries, although these are not required by \textit{SciUnit}.  
To create a \emph{particular} position test, we instantiate this class with particular observations. 
For example, the subset of cosmologists interested specifically in Saturn might instantiate a test by randomly chunking observations made about Saturn's nightly position over time as follows:
\vspace{-3px}
\begin{python}
  h, p = randomly_chunk(saturn_obs_positions)
  saturn_position_test = PositionTest(
    {'histories': h, 'positions': p})
\end{python}
\vspace{-5px}
The class \verb|PositionTest| defines logic that is not specific to any particular planet, so it is contained in a repository shared by all cosmologists called \verbx{cosmounit}. The particular test above would be constructed in a separate repository focused specifically on Saturn called \verbx{saturnsuite}. Both would be collaboratively developed by these (overlapping) research communities in source code repositories on Github.

Classes that implement the \verbx{sciunit.Test} interface must contain a \verbx{generate_prediction} method that takes a \emph{model} as input and extracts a prediction. To specify the interface between the test and the model, the test author specifies a list of \emph{required capabilities} in the \verbx{required_capabilities}, seen on line 7. 
A capability specifies a collection of methods that a test will need to invoke in order to receive relevant data, and is analogous to an \emph{interface} in a language like Java. 
In Python, interfaces are written as classes with unimplemented members. 
The capability required in Figure \ref{fig:rate_test} is shown in Figure \ref{fig:capability}. 
Classes defining capabilities are tagged as such by inheriting from \verbx{sciunit.Capability}. The test in Figure \ref{fig:rate_test} repeatedly uses this capability on lines 9-11 to return a list of predicted positions for each observed history (using a Python list comprehension for brevity).  A planet's position is represented using a standard celestial coordinate system specified within \verbx{cosmounit} by the class \verbx{Position}. A model class can implement a capability by simply inheriting from it and implementing the required methods (Figure \ref{fig:capability}). A particular model is an instance of such a class.  The scope of a model is identified by the capabilities it implements.
\vspace{-3px}
\begin{python}
lin_saturn_model = LinearPlanetModel()
\end{python}
\vspace{-5px}
Once a prediction has been extracted from a model, the test class must compute a \emph{score}. The framework invokes the \verbx{compute_score} method with the observation provided upon test instantiation and the prediction just generated (line 14 of Fig. \ref{fig:rate_test}). Here, the test class constructs a list of distances between the observed and predicted positions, then determines a pooled $p$-value to determine the goodness-of-fit (the details are omitted for concision). A $p$-value is represented as an instance of \verbx{sciunit.PValue}, a subclass of \texttt{sciunit.Score} that has been included with {SciUnit} due to its wide use across science. %In addition to the $p$-value itself, the returned score object also contains metadata, accessible via the \verbx{related_data} attribute, for  scientists who may wish to examine the result in more detail later. This includes the observation, prediction, and a reference to the model and test. 
\begin{figure}
\small
\begin{python}
class PredictsPlanetaryPosition(sciunit.Capability):
  def predict_next_pos(self, history): 
    """Takes a list of previous Positions and produces the next Position."""
    raise NotImplementedError()

class LinearPlanetModel(sciunit.Model, 
    PredictsPlanetaryPosition):
  def predict_next_pos(self, history):
    return linear_prediction(history)
\end{python}
\vspace{-12px}
\caption{\small An example capability and a model class that implements it in \verbx{cosmounit}.}
\label{fig:capability}
\vspace{-10px}
\end{figure}

This illustrates a key difference between unit testing, which would simply produce a boolean result, and our conception of scientific validation testing (challenge 1). A score class must induce an ordering, so that a table like that shown in Figure 1 can be sorted along its columns, and it can  specify a normalization scheme so the cells can be color-coded. Predictions are extracted by a separate method to make it easier for statisticians to write new tests for which only the goodness-of-fit metric differs, not the method by which the prediction is extracted (challenge 5). %It also makes it easier to cache predictions generated by large-scale models so that they can be used by multiple tests.

Te \verbx{judge} method of a test can be invoked to compute a score for a single model: 
\vspace{-4px}
\begin{python}
score = saturn_position_test.judge(lin_saturn_model)
\end{python}
\vspace{-5px}
This method proceeds by first checking that the provided model implements all required capabilities before calling the \verbx{generate_prediction} method followed immediately by the \verbx{compute_score} method. A reference to the test,   model, observation, prediction and other related data the test provides (not shown) are available as attributes of the score.

To produce a more comprehensive test suite, the contributors to \verbx{saturnsuite} would instantiate a number of other tests and then create an instance of the \verbx{TestSuite} class:
\begin{python}
saturn_motion_suite = sciunit.TestSuite([
  saturn_position_test, saturn_velocity_test, ...])
\end{python}
They would also instantiate a number of models. A test suite can be used to judge multiple models at once (challenge 2) if they satisfy the union of the capabilities required by the constituent tests. The result is a \emph{score matrix}:
\begin{python}
sm_matrix = saturn_motion.judge([lin_copernicus_sat, ptolemy_sat, newton_sat, kepler_sat])
\end{python}

If constructed inside an iPython notebook, a score matrix can then be visualized as an interactive table, much like the one shown in Figure \ref{fig:sciunit_overview}. Scientists can sort by column and click on tests, models and scores to get more information on each (not shown, implementation in progress).

%They then defined a combined metric favoring models that broadly succeeded at meeting these criteria, to produce an overall ranking. 
%Such combined criteria are simply validation tests that invoke other tests to produce a result.

@@

@@

%\subsection{Models}
%Capabilities are \emph{implemented} by models. In \textit{SciUnit}, models are instances of Python classes that inherit from \verbx{sciunit.Model}. Like tests, the class itself represents a family of models, parameterized by the arguments of the constructor. A particular model is an instance of such a class.
%
%Figure \ref{fig:simple_model} shows how to write a simple {family} of models, \verbx{LinearModel}, that implement the capability in Fig. \ref{fig:capability} as well as another capability shown in Fig. \ref{fig:training}, which we will discuss below. 
%Models in this family generate a spike count by applying a linear transformation to the mean of the provided input current. The family is parameterized by the scale factor and the offset of the transformation, both scalars. 
%To create a \emph{particular} linear model, a modeler can provide particular parameter values, just as with test families:
%\begin{python}
%CA1_linear_model_heuristic = LinearModel(3.0, 1.0)
%\end{python}
%Here, the parameters to the model were picked by the modeler heuristically, or based on externally-available knowledge. 
%An alternative test design would add a training phase where these parameters were fit to data using the capability shown in Fig. \ref{fig:training}. 
%This test could thus only be used for those models for which parameters can be adjusted without human involvement. 
%Whether to build a training phase into the test protocol is a choice left to each test development community. 
%
%Fig. \ref{fig:rate_test} does not include a training phase. 
%If training data is externally available, models that nevertheless do implement a training capability (like \verb|LinearModel|) can simply be trained explicitly by calling the capability method just like any other Python method:
%\begin{python}
%CA1_linear_model_fit = LinearModel()
%CA1_linear_model_fit.train_with_currents(CA1_training_in, CA1_training_out)
%\end{python}
%\begin{figure}
%\begin{python}
%class LinearModel(sciunit.Model, SpikeCountFromCurrent, 
%    TrainSpikeCountFromCurrent):
%  def __init__(self, scale=None, offset=None): 
%    self.scale, self.offset = scale, offset
%    
%  def spike_count_from_current(self, input):
%    return int(self.scale*numpy.mean(input) + self.offset)
%
%  def train_with_currents(self, currents, counts):
%    means = [numpy.mean(c) for c in currents]
%    [self.offset, self.scale] = numpy.polyfit(means, counts, deg=1)    
%\end{python}
%\caption{A model that returns a spike count by applying a linear transformation to the mean input current. The parameters can be provided manually or learned from data provided by a test or user (see text).}
%\label{fig:simple_model}
%\vspace{-15px}
%\end{figure}
%
\section{Collaborative Workflow}\label{scidash}
A model that performs well across tests in such a suite could reasonably claim (e.g. to reviewers) to be a coherent, valid model of Saturn's motion. As new data is collected, new tests can be added to the suite or existing tests can be refined. As long as the interface between the test and the model remains the same, the table can be updated entirely automatically. Similarly, when a new model is developed, it can immediately be evaluated against all known data that has been encoded as a test by simply exposing its predictions via the capabilities the test requires. If better statistical tests are developed, they too can be used to regenerate the table automatically, as discussed above.

The design just described has been purposefully left as simple as possible, in pursuit of challenge 4. Despite its simple design, it captures the essential elements of the scientific model validation process and satisfies the four challenges we laid out. SciUnit can be used by individual scientists to organize their workflows, but because science, and in particular, model validation is a collaborative process, we have also described the intended use of SciUnit within a collaborative workflow, mediated by a social coding tool like Github.

@@

\section{Discussion}\label{discussion}
Academic peer review and code review are similar, but software developers typically supplement human review with test-driven methodologies. Such methodologies may benefit scientific communities as well, but several unique  constraints make reusing existing testing frameworks and processes difficult. We describe a core testing framework that addresses these issues  by using an existing, widely-adopted language and designing a flexible, simple framework that captures the domain-specific structure of scientific model validation. We then describe, by example, how the various components can be organized into software repositories and collaboratively maintained using social coding tools to support existing scientific practices. We end by outlining a collaboratively filtered portal that sits atop these tools and facilitates exploratory analyses and repository discovery.
